---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---
## Loading and preprocessing the data

```{r, echo = TRUE}
df <- read.csv("activity.csv")

library(ggplot2)
library(plyr)
library(dplyr)
library(chron)

```

## What is mean total number of steps taken per day?

A histogram of the total number of steps taken each day

```{r, echo = TRUE}
c <- ggplot(df, aes(x=date, y = steps))
c + geom_bar(stat = "identity") + theme(axis.text.x = element_text(face = "bold", color = "#993333", size = 5, angle = 90))+labs(x = "Date", y = "Steps", title = " Total number of steps taken each day")
```

The mean and median total number of steps taken per day

```{r, echo = TRUE}
df_mean <- ddply(df,~date,summarise,mean=mean(steps, na.rm = TRUE), median = median 
(steps, na.rm = TRUE))
head(df_mean, 15)
```
## What is the average daily activity pattern?

A time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = TRUE}
df_mean_int <- ddply(df,~interval,summarise,mean=mean(steps, na.rm = TRUE))
plot(df_mean_int$interval, df_mean_int$mean, type = "l", xlab = "Interval", ylab = "Average number of steps", main = "A time series plot of the 5-minute interval")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo = TRUE}
max_int <- subset(df_mean_int, mean ==  max(df_mean_int$mean, na.rm = TRUE), select = interval)
max_int
```
## Imputing missing values

The total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo = TRUE}
max_int <- subset(df_mean_int, mean ==  max(df_mean_int$mean, na.rm = TRUE), select = interval)
max_int
```

## Imputing missing values

The total number of missing values in the dataset is 2034.
```{r, echo = TRUE}
MV <- sum(is.na (df$steps))
MV
```
Devised a strategy for filling in all of the missing values in the dataset. I used the mean for that 5-minute interval.

```{r, echo = TRUE}
NA_find <- function(df) {
  for (i in 1:nrow(df)) {
    if (is.na(df$steps[i])) {
      mean_steps <- mean (subset (df, interval == interval[i], select = steps)$steps, na.rm = TRUE)
      df$steps[i] <- mean_steps 
    }
  }
  df
}
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo = TRUE}
my_df_2 <- NA_find(df)
head(my_df_2, 15)
```

A histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r, echo = TRUE}
df_mean <- ddply(my_df_2,~date,summarise,mean=mean(steps, na.rm = TRUE), median = median (steps, na.rm = TRUE))
c <- ggplot(my_df_2, aes(x=date, y = steps))
c + geom_bar(stat = "identity") + theme(axis.text.x = element_text(face = "bold", color = "#993333", size = 5, angle = 90))+labs(x = "Date", y = "Steps", title = " Total number of steps taken each day")
```

*Do these values differ from the estimates from the first part of the assignment?*

Yes, the NA values are different from the first part of the assignment. 

*What is the impact of imputing missing data on the estimates of the total daily number of steps?*

Now, there are no dates without data about steps . Total average number of steps became higher.

## Are there differences in activity patterns between weekdays and weekends?

A new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo = TRUE}
my_df_2$date <- as.Date(my_df_2$date, format = "%Y-%m-%d")
factor_day <- c ("weekday", "weekend")
factor_day <- factor (factor_day)
my_df_2$day <- 0
my_df_2$day <- as.character(my_df_2$day)
my_df_2$day <- ifelse(is.weekend(my_df_2$date),  my_df_2$day <- c("weekend"),  my_df_2$day <- c("weekday"))
head(my_df_2)
```
A panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r, echo = TRUE}
sub_weekday <- subset(my_df_2, day == "weekday")
mean_weekday <- ddply(sub_weekday,~interval, summarise,mean=mean(steps, na.rm = TRUE))
sub_weekend <- subset(my_df_2, day == "weekend")
mean_weekend <- ddply(sub_weekend,~interval, summarise,mean=mean(steps, na.rm = TRUE))
par(mfrow = c(2,1))
plot(mean_weekday$interval, mean_weekday$mean, type = "l", col = "blue", xlab = "interval", ylab = "average steps", main = "Weekdays")
plot(mean_weekend$interval, mean_weekend$mean, type = "l", col = "red", xlab = "interval", ylab = "average steps", main = "Weekends")
```